---
title: "AI Jobs Remote Work Analysis"
format: html
---

## Introduction

First we need to load the necessary libraries, convert the remote ratio to a factor, handle missing salary values, and then calculate the average salary by remote ratio.

```{r}
library(ggplot2)
library(tidyverse)
library(scales)

# Load and clean data
ai_jobs <- read_csv("../datasets/ai_jobs_dataset.csv")

# Convert remote ratio to descriptive labels
get_factor <- function(x) {
  if (x == 0) {
    return("Completely In-Person")
  } else if (x == 50) {
    return("Hybrid")
  } else {
    return("Fully Remote")
  }
}

# Apply conversion and handle missing salary values
ai_jobs$remote_ratio <- sapply(ai_jobs$remote_ratio, get_factor)
ai_jobs$salary_usd[is.na(ai_jobs$salary_usd)] <- mean(ai_jobs$salary_usd, na.rm = TRUE)

# Calculate average salaries by remote ratio
avg_salary <- ai_jobs %>%
  group_by(remote_ratio) %>%
  summarise(avg_salary_usd = mean(salary_usd))
```

## Data Analysis

Now let's explore the relationship between remote work options and salaries in AI jobs. My hypothesis is that remote jobs pay more than in person jobs.

We ask ourselves, what do we need to do the data to be able to visualize this hypothesis.

So in this case we would want to grab each years spending categories and plot that in something like a vertical bar graph.

```{r}
ggplot(avg_salary, aes(x = remote_ratio, y = avg_salary_usd, fill = remote_ratio)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = dollar) +
  labs(
    title = "AI Job Remote Ratio vs. Average Salary in USD",
    subtitle = "Does working remotely for AI jobs make you more money?",
    x = "Remote Ratio", 
    y = "Average Salary (USD)"
  ) +
  coord_cartesian(ylim = c(112000, 117000)) +
  theme_minimal()
```

## Interpretation

Our analysis shows that our original hypothesis was correct. Remote jobs pay more than in person jobs and hybrid jobs.

1.  Remote jobs pay more than in person jobs and hybrid jobs.
2.  Hybrid jobs pay more than in person jobs but less than remote jobs.
3.  In person jobs pay less than both remote and hybrid jobs.

The data shows that remote jobs pay more than in person jobs. This is expected as remote jobs are often more flexible and allow for a better work-life balance. The data also shows that hybrid jobs pay more than in person jobs but less than remote jobs. This is expected as hybrid jobs allow for a balance between remote and in person work. The data also shows that in person jobs pay less than both remote and hybrid jobs. This is expected as in person jobs are often more expensive to run and require more resources.
